Understanding the Lomb-Scargle Periodogram

### 2017-01-06 PracticalLombScargle.tex
Add multiterm discussion

The paper is organized as follows:
In \sect{continuous-fourier-transform}, we review the continous Fourier
transform and some of its useful properties, including defining the notion of
a power spectrum for measuring periodic content in a signal.
In \sect{window-functions}, we use these properties to understand how different
observation patterns (i.e. window functions) transform the power spectrum, and
discuss a conceptual understanding of the well-known Nyquist limit.
In \sect{non-uniform-sampling}, we consider non-uniformly sampled signals, and
show that the Nyquist limit for this case is quite different than what many
practitioners in the field seem to assume.
In \sect{schuster-to-lomb-scargle}, we discuss the Lomb-Scargle periodogram, a
modified version of the classical periodogram for unevenly-sampled data, as
well as the motivation behind these modifications.
In \sect{practical-considerations}, we discuss important practical
considerations for the use of the Lomb-Scargle periodogram, including
the choice of frequency grid and the various normalizations of the periodogram.
In \sect{uncertainties}, we discuss how to quantify uncertainties in periods
derived from the periodogram, including various notions of the
``False Alarm Probability''.
In \sect{generalizations}, we consider and critique some generalizations of
the Lomb-Scargle model, including multiterm periodograms and Bayesian
approaches.

@@

A further generalization involves multi-term Fourier models at each frequency:
rather than fitting just a single sinusoid at each frequency, we might fit a
partial Fourier series at each frequency, adding $K$ additional components at
integer multiples of the fundamental frequency:
\begin{equation}
  y_{model}(t;f) = y^{0}_f + \sum_{k=1}^K A_f^{(k)} \sin(2\pi k f (t - \phi_f^{(k)})).
\end{equation}
This type of multi-term Lomb-Scargle periodogram was studied in detail by
\citet{Bretthorst88}.

This kind of Fourier series generalization to the periodogram is quite
tempting, because it means that the periodogram can be tuned to fit models
that are more complicated than simple sine waves.
\citet{Bretthorst88} takes this even further, exploring applications to
decaying signals and other more complex models.

@@

\begin{figure}[ht]
  \centering
  \includegraphics[width=\textwidth]{fig21_kepler_multiterm}
  \caption{Multi-term models (left) and their corresponding periodograms (right)
    for the kepler data shown in \fig{kepler-data}.
    \figlabel{kepler-multiterm}}
\end{figure}

@@

>>>>

The paper is organized as follows:
In \sect{continuous-fourier-transform}, we review the continous Fourier
transform and some of its useful properties, including defining the notion of
a power spectrum for measuring periodic content in a signal.
In \sect{window-functions}, we use these properties to understand how different
observation patterns (i.e. window functions) transform the power spectrum, and
discuss a conceptual understanding of the well-known Nyquist limit.
In \sect{non-uniform-sampling}, we consider non-uniformly sampled signals, and
show that the Nyquist limit for this case is quite different than what many
practitioners in the field seem to assume.
In \sect{schuster-to-lomb-scargle}, we discuss the Lomb-Scargle periodogram, a
modified version of the classical periodogram for unevenly-sampled data, as
well as the motivation behind these modifications.
In \sect{practical-considerations}, we discuss important practical
considerations for the use of the Lomb-Scargle periodogram, including
the choice of frequency grid and the various normalizations of the periodogram.
In \sect{uncertainties}, we discuss how to quantify uncertainties in periods
derived from the periodogram, including various notions of the
``False Alarm Probability''.

@@

A further generalization involves multi-term Fourier models at each frequency:
rather than fitting just a single sinusoid at each frequency, we might fit a
partial Fourier series at each frequency, adding $K$ additional components at
integer multiples of the fundamental frequency:
\begin{equation}
  y_{model}(t;f) = y^{0}_f + \sum_{k=1}^K A_f^{(k)} \sin(2\pi k f (t - \phi_f^{(k)})).
\end{equation}
\citet{Bretthorst88} takes a comprehensive look at this type of multi-term
extension to the periodogram, as well as extensions to decaying signals and
other more complex models.

@@

This kind of Fourier series generalization to the periodogram is quite
tempting, because it means that the periodogram can be tuned to fit models
that are more complicated than simple sine waves.
In some cases, this can be very useful; for example, \fig{binary-multiterm}
shows a Lomb-Scargle analysis of an eclipsing binary star, characterized by
both a primary and secondary eclipse.
The standard Lomb-Scargle periodogram (upper panels) is maximized at twice the true rotation
frequency, because the simple sinusoidal model is unable to closely fit the
primary and secondary eclipses separately.
A six-term Fourier model (lower panels) is sufficiently flexible that it can
detect the true 17.5-hour period, though this comes at the expense of a much
noisier periodogram.

This background noise is easy to understand: in the least-squares view of the
periodogram, the periodogram height at any frequency is directly related to
how well the model fits the data.
For Linear models, adding additional terms will {\it always} provide an
equal or better fit to data than the simpler model, and so a periodogram based
on a more complex model will be higher {\it at all frequencies}, not just
at frequencies of interest!

\begin{figure}[ht]
  \centering
  \includegraphics[width=\textwidth]{fig21_kepler_multiterm}
  \caption{Multi-term models (left) and their corresponding periodograms (right)
    for the kepler data shown in \fig{kepler-data}
    \figlabel{kepler-multiterm}}
\end{figure}

This background noise can also be understood in terms of aliasing.
Consider \fig{kepler-multiterm}, which shows several multi-term fits to the
data from \fig{kepler-data}.
Given a standard periodogram peak at $f_0$, a 2-term periodogram will have an
additional peak at $f_0 / 2$, in which the original peak is the second
harmonic. Similarly, a 3-term periodogram will add additional peaks both at
$f_0 / 2$ and $f_0 / 3$, due to the original peak falling in the second and
third harmonic.

@@

In cases where the single-term power spectrum is itself very noisy, these
additional aliases can quickly wash-out any gains from the more complicated
model.
Consider \fig{LINEAR-multiterm}, in which the $n=3$ model leads to very large
noise at the low-frequency end. In this case, spurious periodogram peaks due
to coincidences of such aliases may become problematic for even moderately
noisy models.
